\chapter{State of the Art}
\label{ch:State_of_the_Art}

Motorised road transport emitted nearly \SI{8}{\giga\tonne} of CO\textsubscript{2} in 2022—accounting for over one-third of all end-use emissions. \cite{IEATransport2024}  
Reducing these emissions effectively requires an in-depth understanding of traffic dynamics and energy consumption at the individual vehicle trajectory level.   
Contemporary research thus integrates  
\begin{enumerate}[label=(\roman*)]
  \item high-fidelity microscopic traffic simulation,
  \item optimisation frameworks such as \ac{glosa},
  \item data-driven or learning-based control strategies.
\end{enumerate}
This chapter reviews the fundamental components of that pipeline. Section~\ref{sec:SUMO} introduces \ac{sumo}, the open-source simulation platform employed throughout this thesis. Section~\ref{sec:glosa} describes the underlying principles and system architecture of \ac{glosa}.
\mynewline
The subsequent discussion is organised as follows. First, we present \ac{dp}–based approaches aimed at maximising traffic throughput (Section~\ref{sec:flow_glosa}). These methods compute optimal speed profiles that minimise stop‐and‐go behaviour and thus increase network capacity. Next, we extend the \ac{dp} framework to incorporate energy considerations, formulating an \ac{eco-glosa} algorithm (Section~\ref{sec:eco_glosa}) that prioritises minimising fuel consumption through trajectory-level speed optimisation while still enhancing traffic throughput. Finally, we turn to \ac{rl}–based techniques (Section~\ref{sec:rl_eco_glosa}), where a learning agent derives adaptive speed advisories from real‐time traffic states to further enhance both efficiency and sustainability.
\mynewline
Section~\ref{sec:research_gap} then synthesises the remaining research gaps that motivate the contributions of this work.  

\input{chapters/02-StateOfTheArt/021-SUMO}

\input{chapters/02-StateOfTheArt/022-GLOSA}

\input{chapters/02-StateOfTheArt/023-DynamicProgrammingForSpeedOptimization}

\input{chapters/02-StateOfTheArt/024-EcoDrivingAndFuelConsumptionReduction}

\input{chapters/02-StateOfTheArt/025-ReinforcementLearningInTrafficApplications}

\input{chapters/02-StateOfTheArt/026-ResearchGapsAndConclusions}