\section{\acl{dp} for Energy-Aware \acs{glosa}}
\label{sec:eco_glosa}

\ac{glosa} algorithms advise road users on speed adjustments that maximise the probability of crossing signalised intersections during green phases. When the objective is to minimise energy use rather than merely reduce delay, \ac{dp} offers a mathematically rigorous means of embedding high-fidelity powertrain models, signal timing constraints, and safety limits into a single optimal-control framework. The appeal of \ac{dp} lies in its ability to guarantee local optimality over a discretised state-action space, to accommodate non-convex efficiency maps and integer variables, such as gear or hybrid mode selection, and to incorporate hard constraints like arrival windows or legally mandated speed bounds.
\mynewline
An energy-conscious \ac{glosa} \ac{dp} must address three fundamental modelling questions. First, for the \textit{objective formulation}, does the cost functional target sole fuel minimisation, or does it balance energy with other factors like delay, comfort, or even pollutant exposure? Second, concerning the \textit{powertrain representation}, should the stage cost rely on tractive-energy proxies, such as mass-speed-acceleration products, or on calibrated maps like PHEMlight5, HBEFA4, or VT-CPFM, which convert wheel demands into fuel mass and \ac{co2} in grams? Third, for the \textit{constraint encoding}, how are vehicle dynamics (speed, longitudinal acceleration, jerk), legal limits, and powertrain ceilings (engine torque, battery state-of-charge, electric motor power) enforced without causing the dimension of the state space to explode? The following subsection synthesises how recent studies have tackled these three questions, highlighting common design patterns, trade-offs, and emerging trends toward multi-criteria optimisation.

\subsection{Energy and Multi-Criteria Objective Formulations}
\label{subsec:eco_dp_objectives}

An \ac{eco-glosa} controller seeks to find a longitudinal control signal, $u(t)$, which is typically a bounded acceleration or wheel torque. The resulting vehicle trajectory should minimise an energy-centric performance index while guaranteeing legal and safety compliance at signalised intersections. The system state can be represented by the vector $\mathbf x(t) = [s(t), v(t), \mathrm{SoC}(t)]^{\!\top}$, which comprises the travelled distance $s$, speed $v$, and, for electrified powertrains, the battery's state of charge. The continuous-time dynamics can be abstracted as:
\begin{equation*}
    \dot{\mathbf x}(t) = \mathbf f\bigl(\mathbf x(t), u(t)\bigr),
    \qquad
    u_{\min} \le u(t) \le u_{\max},
\end{equation*}
where $\mathbf f$ captures the longitudinal equations of motion, rolling and aerodynamic resistances, grade, and a simple battery equivalence for hybrids. Discretising this formulation on a uniform temporal grid, $t_k = k \Delta t$, yields the canonical \ac{dp} recursion:
\begin{equation*}
    V_k(\mathbf x_k) = \min_{u_k \in \mathcal U} \left\{ g_k(\mathbf x_k, u_k) + V_{k+1}\bigl(\mathbf x_{k+1} \bigr) \right\},
    \qquad
    \mathbf x_{k+1} = F_k(\mathbf x_k, u_k).
\end{equation*}
This recursion is subject to a terminal cost, $V_{N}(\mathbf x_N)=\Phi(\mathbf x_N)$, which encodes the mandatory arrival time $t_f$, the stop-bar position $s_f$, and the permissible speed range $v\in[v_{\min},v_{\max}]$ at the intersection.

\paragraph{Single-Criterion Energy Minimisation.}
The most direct objective is the integral of instantaneous fuel or electrical power $P_{\text{tr}}$:
\begin{equation*}
\label{eq:single}
    J_{\text{fuel}} = \int_{0}^{t_f} P_{\text{tr}}\bigl(\mathbf x(t), u(t)\bigr)\,\mathrm{d}t
    \quad\Longrightarrow\quad
    g_k = \Delta t \cdot P_{\text{tr}}\bigl(\mathbf x_k, u_k\bigr).
\end{equation*}
For conventional engines, $P_{\text{tr}}$ is mapped to mass flow $\dot m_{\text{fuel}}$ via calibrated look-up tables such as PHEMlight5 or HBEFA4. For battery-electric vehicles it equals traction power minus regenerative recovery, and for hybrids it combines engine fuel and battery energy through an equivalence factor $\xi$ so that $g_k = \Delta t \cdot \left[ \dot m_{\text{fuel}} + \xi P_{\text{bat}} \right]$.

\paragraph{Weighted-Sum Multi-Objective Costs.}
Minimising Equation~\eqref{eq:single} alone can drive the optimum toward unrealistically low speeds. A pragmatic remedy is to augment the single-criterion formula with travel time $T = t_f - t_0$ and comfort penalties on jerk $j = \dot a$:
\begin{equation*}
    J_{\mathrm{sum}} = \alpha E + \beta T + \gamma \int_{0}^{t_f} j(t)^2\,\mathrm{d}t,
    \qquad
    E = \int_{0}^{t_f} P_{\text{tr}}\,\mathrm{d}t,
\label{eq:weighted}
\end{equation*}
where $\alpha$, $\beta$, and $\gamma$ tune the policy along the energy–delay–comfort trade-off surface. In \ac{dp} form, the stage cost becomes
$g_k = \alpha \Delta t\,P_{\text{tr}} + \beta \Delta t + \gamma \Delta t\,j_k^2$.

\paragraph{Pareto Formulations.}
To avoid arbitrary weights, an \emph{$\epsilon$-constraint} variant minimises one metric while bounding the others:
\begin{equation*}
    \min\;E \;\;\text{s.\,t.}\;\;T \le \bar T, \qquad
    g_k = \Delta t\,P_{\text{tr}}, \qquad
    \Phi = \mathbbm{1}_{T \le \bar T},
\end{equation*}
generating Pareto-optimal fronts by sweeping the tolerance $\bar T$. A higher-level supervisory module can then select the point that best fits network objectives or driver preference.

\paragraph{Constraint Set.}
All formulations inherit hard bounds from road regulations and power-train physics:
\begin{align*}
    0 &\le v_k\le v_{\mathrm{lim}}(s_k), &
    |a_k| &\le a_{\max}, &
    |j_k| &\le j_{\max}, \nonumber\\
    P_{\text{eng},k} &\le P_{\text{eng}}^{\max}, &
    \mathrm{SoC}_{\min} &\le \mathrm{SoC}_k\le \mathrm{SoC}_{\max}, &
    t_{\mathrm{g,\,on}}\le t_k &\le t_{\mathrm{g,\,off}},
\end{align*}
where the final line enforces the \ac{spat}-derived arrival window. Constraints are handled either by pruning infeasible lattice nodes before the Bellman sweep or by adding large penalties to $g_k$.

\paragraph{Lattice Design and Complexity Control.}
Let $n_s$, $n_v$ and $n_{\mathrm{SoC}}$ be the discretisation levels of distance, speed, and battery charge. The complexity of the Bellman formula scales as $\mathcal O(N\,n_s n_v n_{\mathrm{SoC}} N_u)$, with $N_u$ the number of admissible controls. To render on-board computation feasible, three reduction strategies dominate:  
(1)~\textit{Variable Grids} contract the lattice where the value function is nearly affine,  
(2)~\textit{Multi-Stage} splitting solves the Bellman formula separately across signal phases or queue discharge events and stitches the partial value functions at boundaries, and  
(3)~\textit{Surrogate Models} replace costly $P_{\text{tr}}$ evaluations with neural or polynomial emulators trained offline.
\mynewline
The above framework, which includes a rigorous Bellman recursion, a cost functional that can be single or multi-objective, and systematic constraint handling, forms the basis for energy-aware \ac{glosa} speed guidance. Subsequent subsections build on this scaffold to incorporate stochastic queues, \ac{v2x} connectivity and cooperative manoeuvres.

\subsection{Queue–Connectivity–Aware \ac{eco-glosa}}
\label{subsec:eco_dp_queue}

This section reviews eight \ac{dp}-based eco-driving strategies that are designed to minimise fuel or electrical energy consumption while respecting signal timing and queue dynamics. The discussion begins with single-vehicle, offline \ac{dp} formulations for battery and hybrid powertrains, as explored by Park et al. \cite{Park2024} and Pulvirenti et al. \cite{Pulvirenti2023}. It then proceeds to cover real-time, multi-stage, and queue-aware controllers from researchers such as Kamalanathsharma et al. \cite{Kamalanathsharma2013} and He et al. \cite{He2015}. The section concludes with an examination of cooperative, \ac{v2i}/\ac{v2v}-enabled schemes developed by Yang \cite{Yang2017}, Ala \cite{Ala2016}, Yang \cite{Yang2021}, and Dong \cite{Dong2024}. For each method, the \ac{dp} formulation is summarised, precise savings and runtimes are reported, and key limitations and suggested remedies are highlighted.
\mynewline
To begin, Park et al. \cite{Park2024} tackle the offline \ac{bev} case by formulating an eco-driving problem for a Hyundai Ioniq 5. They use a two-state (distance, speed) discrete-time \ac{dp} model that minimises cumulative electrical power. This is achieved by exhaustively enumerating acceleration commands on a $1\unit{\second}$ grid while enforcing fixed departure and arrival times, distance and speed limits, and drivetrain comfort bounds. The solver explores a $150\unit{\second}$ horizon with a $0.05\unit{\metre}$ space step and a $0.1\unit{\metre\per\second}$ speed step, which incurs $337\unit{\second}$ of offline computation. Comparing the \ac{dp} trajectories with an \ac{idm} baseline shows sizeable battery savings. Over a $500\unit{\metre}$ sprint, a \enquote{Comfort-DP} profile with human-like acceleration cuts energy use by $34.3\%$, while an \enquote{Aggressive-DP} profile using full torque reduces it by $67.5\%$. Over a $900\unit{\metre}$ run, the respective reductions are $25.5\%$ and $47.8\%$. The analysis attributes these gains to two factors: the strict use of high-efficiency tangential and stationary points on the motor's Pareto frontier during acceleration and cruising, and full energy recuperation during terminal braking. The limitations of this approach include its reliance on perfect prior knowledge of boundary conditions and zero road grade, the absence of battery power limits at low \ac{soc}, and infeasible on-board runtimes without a GPU or surrogate policies. The authors, therefore, advocate for state-reduction heuristics, stochastic \ac{spat} and traffic integration, and real-vehicle validation to bridge the gap between offline optimality and real-time deployment.
\mynewline
While Park et al. focus on an exhaustive grid search for \acp{bev} under fixed boundary conditions, Pulvirenti et al. \cite{Pulvirenti2023} extend the \ac{dp} paradigm to plug-in hybrids by introducing a \ac{vgdp} framework. This framework exploits \ac{v2x} data and cloud computation to generate energy-optimal speed trajectories for a plug-in hybrid Mercedes E300de along a $96\unit{\kilo\metre}$ RDE-compliant route. The spatial-domain \ac{dp} state is defined as either $(v)$ for a stop-controlled scenario or $(v,t)$ for a traffic-light scenario, with longitudinal acceleration as the control variable. The bi-objective stage cost combines incremental traction energy with travel time via a weighting factor $\beta$. Instead of searching a fixed lattice, \ac{vgdp} shrinks the speed grid to an interval defined by a $500\unit{\metre}$ moving-average envelope ($\pm20\unit{\kilo\metre\per\hour}$) and bounds the time grid around an ETA corridor derived from communicated \ac{spat} data, speed limits, and grades. With a $5\unit{\metre}$ spatial step, this method reduces the node count by two orders of magnitude and slashes the solver wall-time on an Intel E5-2680 workstation from $24\unit{\minute}$ to under $1\unit{\minute}$ (a $95\%$ reduction), without degrading optimality. In \textit{Scenario 1}, with stop signs imposed at every observed standstill, the optimiser delivers smoother acceleration profiles that lower the fuel-equivalent energy by $29\%$ and shrink the cruise time by $10\%$, while still respecting all full-stop constraints. In \textit{Scenario 2}, all stops are replaced by traffic lights with a $60\unit{\second}$ cycle and $60\%$ green time, where the \ac{spat} data is assumed to be deterministic and known. Here, \ac{vgdp} schedules an almost uniform $49\unit{\kilo\metre\per\hour}$ cruise, eliminates complete halts, and achieves $54\%$ energy and $38\%$ travel-time savings relative to the reference drive. A sensitivity analysis shows only modest Pareto trade-offs as $\beta$ varies from $0.2$ to $0.8$. The limitations include the reliance on perfect \ac{spat} and queue-free flow, a single-vehicle scope, and the need for a high-resolution distance discretisation ($5\unit{m}$), which still burdens memory when scaled to network traffic. Future work is therefore directed at platoon-level coordination, stochastic queue integration, and on-road hardware validation of the cloud-assisted control loop.
\mynewline
Moving from cloud-assisted, offline trajectory planning to real-time control, Kamalanathsharma et al. \cite{Kamalanathsharma2013} propose a \textit{\ac{msdp} eco-speed controller}. This controller frames the approach to a red phase as a least-cost path-finding problem on a two-stage lattice. The upstream stage chooses a constant deceleration, $d$, and a cruise time, $t_{\mathrm c}$, that brings the vehicle to the stop-bar exactly at the green onset. The downstream stage then selects a throttle trajectory to re-accelerate the vehicle to its desired speed. Both stages minimise the cumulative fuel consumption, as predicted by the VT-CPFM model, subject to comfort bounds ($d\le3\unit{\metre\per \second\squared}$, $a^{+}\le1.1\unit{\metre\per\second\squared}$), grade, weather, and microscopic resistance forces. A heuristic $\mathcal{A}$-star recursion evaluates each $0.1\unit{\second}$ step, pruning dominated nodes so that real-time feasibility is maintained despite the use of full powertrain and aerodynamics equations.
Agent-based MATLAB simulations over a $200\unit{\metre}$ upstream and $400\unit{\metre}$ downstream zone, with random arrivals across an $84\unit{\second}$ cycle and eighteen combinations of speed limits $(25/35/45\unit{\mile\per\hour} \approx 40/56/72\unit{\kilo\metre\per\hour})$, pavement conditions (dry, wet, snow), and grades ($\pm5\%$), report mean savings of $19.5\%$ in fuel and $32\%$ in travel time relative to an ITE-calibrated human baseline. The best-case results reached $37.2\%$ and $41.5\%$ under snowy, $25\unit{\mile\per\hour} \approx 40\unit{\kilo\metre\per\hour}$) conditions. Car-following tests that couple the Rakha–Pasumarthy–Adjerid model show that a non-instrumented follower inherits an approximate $15\%$ fuel reduction when tracking an MS-DP lead vehicle, which indicates spill-over benefits. The limitations of this approach are its reliance on perfect \ac{spat}, a single-vehicle optimisation that ignores queue spill back, sensitivity to lane-changing below a $30\%$ \ac{cv} penetration rate, and validation being confined to simulation. Oversaturated demand, at or above $800\unit{\veh\per\hour}$, breaks the kinematic-wave predictor and can render the benefits negative. The authors therefore call for \ac{v2v}-aided queue sensing, speed-harmonisation overlays, multi-intersection coordination, robustness to packet loss and driver non-compliance, and field trials to confirm the observed potential for approximately $20\%$ fuel savings.
\mynewline
In a similar real-time vein, He et al. \cite{He2015} formulate a \ac{qost} that embeds stochastic queue-tail constraints and non-blocking requirements into a multi-stage optimal-control problem, which enables sub-second computation. In this manner, a single car never impedes traffic moving upstream and does not cause the green light to go out for other vehicles. A five-variable quadratic surrogate, which includes terminal time, two accelerations, and two switch instants, solves each stage in approximately $1.5\unit{\second}$ on off-the-shelf hardware, making real-time deployment plausible. In a two-intersection benchmark, \ac{qost} eliminates a five-second idle time at the first queue. However, because it cruises at an eco-speed of $30\unit{\mile\per\hour} \approx 48\unit{\kilo\metre\per\hour}$, the run time grows from $77.3\unit{\second}$ without advice to $90.0\unit{\second}$, an increase of $16\%$. Simultaneously, stops fall to zero and fuel use drops by $41\%$. A six-signal field trial on Minneapolis TH-55 repeats this pattern. The advised vehicle traverses the corridor without stopping, yet its travel time rises from $236\unit{\second}$ to $258\unit{\second}$ ($+9\%$) as queues are absorbed upstream. This results in a $29\%$ fuel saving and a net fuel-economy improvement from $17\unit{\mile\per\gallon} \approx 13.8\unit{\liter\per100\kilo\metre}$ to $24\unit{\mile\per\gallon} \approx 9.8\unit{\liter\per100\kilo\metre}$.
By design, the model also compels the subject vehicle to clear the stop-line before a threshold, $T_{\tilde B}$, so as not to lengthen the residual queue or trap following traffic in the next red phase, thereby avoiding secondary delay propagation. The authors acknowledge that the traffic-flow benefits are therefore mixed. While idling, shockwaves, and spill back are curtailed, the mean travel time for the advised vehicle, and hence the point throughput, can worsen unless the eco-speed is increased or applied jointly to a platoon. Key challenges include (i) a reliance on high-fidelity, real-time queue estimation, where biases can jeopardise both feasibility and flow gains; (ii) piecewise-constant accelerations that introduce jerk peaks; and (iii) a single-vehicle scope, which cannot exploit cooperative gaps or optimise a multi-vehicle order. Proposed remedies include embedding stochastic queue predictors, smoothing trajectories with powertrain-aware penalties, extending the formulation to the simultaneous optimisation of several vehicles at the front of the queue, and implementing rolling re-optimisation when signals skip phases. These are all aimed at transforming the current fuel-centric design into one that also boosts corridor-level throughput under varying demand patterns.
\mynewline
Building on single-vehicle queue awareness, Yang et al. \cite{Yang2017} introduce a \textit{queue-aware Eco-Cooperative Adaptive Cruise Control} (Eco-CACC-Q) method. This is a cooperative, \ac{v2i}-enabled adaptive cruise control system that uses \ac{spat} packets and kinematic-wave queue estimates to optimize fuel consumption across a platoon. Using this information, advisory speed limits are calculated and updated every second to minimize the instantaneous VT-CPFM fuel rate and to ensure that the probe vehicle reaches the stop-bar precisely when the final vehicle in the queue is released. A receding-horizon quadratic program simultaneously selects an upstream constant deceleration, $a^{-}$, a cruise speed, $v_{\mathrm c}$, over the remaining approach, and a downstream acceleration, $a^{+}$, on a $500\unit{\metre}/200\unit{\metre}$ control segment. The search is bounded by comfort limits of $0\le a^{-}\le3\unit{\metre\per \second\squared}$ and $0\le a^{+}\le2.5\unit{\metre\per\second\squared}$, as well as by \ac{soc}-neutrality constraints.
Single-lane integration simulations with a $20\%$ \ac{mpr} and a uniform $500\unit{\veh\per\hour}$ inflow show that Eco-CACC-Q trims the average fuel use by $11.4\%$ relative to the base car-following model and by $4.5\%$ compared with the non-queue Eco-CACC variant. It also eliminates complete stops and cuts the speed variance from $29.7$ to $15.3\unit{\kilo\metre\per\hour}$. The savings scale almost linearly with penetration, reaching $18.0\%$ at $100\%$ \ac{mpr}. On two-lane approaches, the algorithm is beneficial only above a $30\%$ \ac{mpr}. Below that threshold, aggressive lane changes around slower controlled vehicles inflate acceleration spikes and raise network fuel use by approximately $5\%$. At full penetration, the multi-lane test bed records an $18.3\%$ overall saving, with a $19.2\%$ saving for the \ac{cacc} fleet itself. Key limitations include a dependence on accurate queue-length prediction, as errors from lane-changing or oversaturation negate the benefits. Other limitations include audio or control-actuation latency and the assumption of isolated intersections. At very high demand, the method fails once the LWR-predicted queue exceeds the control horizon. The authors, therefore, call for \ac{v2v}-aided queue sensing, corridor-level coordination with speed-harmonisation controllers, robustness analyses against packet loss and driver non-compliance, and real-vehicle validation to confirm the simulation gains.
\mynewline
Complementing the algorithmic development, Ala et al. \cite{Ala2016} perform an extensive INTEGRATION-based sensitivity study of Eco-CACC-Q, quantifying its performance over varying penetration rates, green splits, and demand levels. Although their contribution is a broad sensitivity study rather than a new controller, they evaluate the Eco-CACC-Q algorithm, which exploits \ac{v2i} \ac{spat} packets and kinematic-wave queue estimates to compute an advisory trajectory each second. This trajectory is composed of a bounded upstream deceleration, $a^{-}\in[0,3]\unit{\metre\per\second\squared}$, a cruise speed, $v_{\mathrm c}$, timed to reach the stop-bar just as the queue clears, and a downstream re-acceleration, $a^{+}\le 2.5\unit{\metre\per\second\squared}$.
Single-lane tests with a fixed $500\unit{\metre}$ upstream and $200\unit{\metre}$ downstream control horizon and a demand of $300\unit{\veh\per\hour}$ show fuel-consumption savings that scale almost linearly with the \ac{mpr}, peaking at $19\%$ when the \ac{mpr} is $100\%$. On two-lane approaches, the algorithm yields negative benefits below a $30\%$ \ac{mpr} because lane-changing around slower controlled vehicles injects oscillations. Above that threshold, the savings recover and reach $19\%$ at full penetration. Varying the green split from $0.3$ to $0.7$ of an $84\unit{\second}$ cycle changes the baseline fuel use but alters the Eco-CACC-Q savings by at most $2\%$. Increasing the upstream control length from $200\unit{\metre}$ to $700\unit{\metre}$ boosts single-lane gains from $12\%$ to $18\%$, with marginal returns beyond $500\unit{\metre}$. The same pattern holds for two-lane links once the \ac{mpr} is greater than $30\%$. The demand sensitivity analysis reveals an optimal flow of $500\unit{\veh\per\hour}$ for a single lane and $600\unit{\veh\per\hour}$ for two lanes. Oversaturated flow, at $800\unit{\veh\per\hour}$, breaks the kinematic-wave queue predictor, causing rolling queues and stop-and-go waves, which turns the benefits negative. A four-leg junction case with asymmetric volumes of $1000\unit{\veh\per\hour}$ on the major road confirms the need for high penetration. Below a $25\%$ \ac{mpr}, fuel use rises, but at full penetration, network consumption drops by up to $25\%$. The limitations highlighted include: (i) a dependence on accurate real-time queue length, as errors or rolling queues undermine the trajectory; (ii) negative effects at low \ac{mpr} on multilane arterials due to lane-changing; (iii) algorithm breakdown under oversaturation; (iv) the assumption of isolated intersections and deterministic \ac{spat}; and (v) the absence of field validation. Future work should therefore target \ac{v2v}-aided queue sensing, speed-harmonisation overlays to throttle entry flow, multi-intersection coordination, and robustness analyses against packet loss and driver non-compliance.
\mynewline
Extending the approach to multi-signal corridors, Yang et al. \cite{Yang2021} propose \ac{ecomsq}, which is a modular, queue-aware, multi-signal controller that broadcasts second-by-second advisory speed limits to \acp{cv}. Leveraging \ac{v2i} \ac{spat} packets and kinematic-wave queue estimates, the algorithm solves a three-segment optimal-control problem for each approaching vehicle. This involves: (i) decelerating at a bounded rate of $a^{-}\in[0,3]\unit{\metre\per\second\squared}$ to a cruise speed, $v_{\mathrm c,1}$, that ensures the stop-bar is reached exactly as the last queued vehicle departs; (ii) re-accelerating or further decelerating to a second cruise speed, $v_{\mathrm c,2}$, that aligns with the downstream signal; and (iii) accelerating at $a^{+}\le 2.5\unit{\metre\per\second\squared}$ back to the link speed limit after clearing the final queue, all while maintaining \ac{soc} neutrality for the $48\unit{\volt}$ mild-hybrid powertrain model.
Implementation in the INTEGRATION micro-simulator shows that on a two-signal arterial with a demand of $600\unit{\veh\per\hour}$, fuel use falls by $7.0\%$ at $100\%$ \ac{mpr}, compared to $4.2\%$ for single-signal control, and the per-vehicle speed variance drops by $30\%$. Extending the analysis to a four-signal corridor with $600\unit{\metre}$ spacing yields savings of $7.7\%$ for single-lane links and $4.8\%$ for two-lane links. A sixteen-intersection grid posts up to a $15\%$ system-wide reduction, with benefits persisting for all \acp{mpr} because side-street and left-turn movements are single-lane constrained. A sensitivity analysis identifies several key factors: (a) an optimum demand of $400\unit{\veh\per\hour}$, which saves $13.5\%$; (b) short green splits ($35\%$ of a $120\unit{\second}$ cycle), which raise savings to $13.8\%$; (c) suboptimal offsets (e.g., $100\unit{\second}$ versus the $45\unit{\second}$ green-wave optimum), which boost savings to $13.0\%$; and (d) a $700\unit{\metre}$ signal spacing, which gives a peak improvement of $13.1\%$. Conversely, offsets near the green-wave optimum cut gains to $2.8\%$. Below a $30\%$ \ac{mpr} on multi-lane arterials, lane-changing around slower \acp{cv} can negate benefits, turning savings negative until the cooperative density exceeds this threshold. Under over-saturated demand ($1000\unit{\veh\per\hour}$), rolling queues breach the LWR predictor, which slashes savings to just $2.7\%$ even at a $10\%$ \ac{mpr}. The limitations, therefore, include a dependence on accurate real-time queue length and dissipation forecasts, as errors propagate into suboptimal cruise speed choices. Other limitations are sensitivity to lane-changing when the \ac{mpr} is low, diminished efficacy near optimal offsets or long green phases, and a lack of spill back handling in over-saturation. The authors call for \ac{v2v}-aided queue sensing, stochastic queue models, corridor-level coordination, and adaptive speed-harmonisation overlays to maintain benefits under high load and mixed traffic.
\mynewline
Finally, Dong et al. \cite{Dong2024} integrate overtaking manoeuvres into the eco-approach control by coupling lane selection via \ac{dp} with a Pontryagin-based speed optimizer, thereby achieving combined lane-planning and energy savings. Using a two-stage receding-horizon framework, the \textit{overtaking-enabled eco-approach control} (OEAC) method yields closed-form control torques while respecting safety headways, comfort bounds ($a^{-}\in[0,3]\unit{\metre\per\second\squared}$, $a^{+}\le2.5\unit{\metre\per\second\squared}$), and deterministic \ac{spat} constraints. The first stage casts the lane-selection task as a finite-horizon Markov decision process, which is solved by \ac{dp} and explicitly models disturbances from surrounding vehicles.
Extensive Monte-Carlo simulations over $10,000$ randomised urban scenarios show that the OEAC method cuts the composite \enquote{monetised} driving cost by an average of $20.91\%$ versus a constant-speed (CS) car-following baseline and by $5.62\%$ versus a regular eco-approach (READ) controller, with maxima of $59.97\%$ and $59.02\%$, respectively. In a representative moderate-flow case, with a vehicle density of $120\unit{\veh\per\kilo\metre}$, the OEAC executes a single lane change, which enables passage in the first green window. This cuts the travel time by $55.93\%$, traction energy by $36.43\%$, and the total cost by $54.12\%$ relative to the CS baseline. Compared to the READ controller, the reductions were $55.89\%$, $31.46\%$, and $53.01\%$, respectively. The computational overhead is negligible, with a mean step time of $0.2\unit{\milli\second}$ and a worst-case time of $9.3\unit{\milli\second}$ on an i9-12900K PC, which is well below the $10\unit{\milli\second}$ simulation tick. The limitations include a dependence on accurate, real-time queue-length and \ac{spat} data, as errors or oversaturation can negate the benefits. Other limitations are the reversion to pure car-following when safe overtakes are unavailable, negative fuel savings below a $30\%$ \ac{mpr} on multilane links due to cut-ins, and the validation being restricted to simulation.
\mynewline
The surveyed \ac{dp}-based \ac{glosa} methods span a range from offline, single-vehicle formulations for \acp{bev} and hybrids to real-time controllers and cooperative \ac{v2x}-enabled schemes. Offline \ac{dp} approaches, such as those by Park et al. and Pulvirenti et al., achieve substantial energy savings, which can be up to $67.5\%$ in \acp{bev} and $54\%$ for hybrids, but they do so at the cost of high computation time or a reliance on cloud resources. Real-time \ac{msdp} and \ac{qost} frameworks, developed by researchers like Kamalanathsharma et al. and He et al., reduce the run-time to sub-second levels while still maintaining fuel or energy reductions of $20\%$--$40\%$ by embedding queue and non-blocking constraints. Cooperative eco-cruise controllers, including those by Yang (2017), Ala (2016), and Yang (2021), leverage \ac{spat} data and kinematic-wave models to achieve fleet-level savings of $7\%$--$19\%$, depending on \ac{mpr}, intersection spacing, and signal timing. The OEAC framework by Dong et al. further incorporates lane-change decisions to realise composite cost reductions of up to $21\%$.
Despite these advances, common limitations persist across the literature. These include assumptions of perfect \ac{spat} and queue knowledge, a scope that is often limited to single vehicles or platoons, sensitivity to low penetration rates and oversaturated demand, and a predominance of simulation-based validation. Future research should therefore target stochastic \ac{spat} and queue modelling, the use of surrogate policies or machine learning for state reduction to enable on-board, real-time deployment, and multi-intersection and network-level coordination. Furthermore, robustness to communication latency and packet loss needs to be addressed, and extensive real-world experiments are required to validate and refine these \ac{dp}-based eco-driving strategies.