\section{Dynamic Programming for Throughput-Oriented GLOSA}
\label{sec:flow_glosa}

\subsection{Objectives and Problem Formulation}
\label{subsec:flow_dp_formulation}
\begin{itemize}
    \item Objective: Minimize cumulative delay, stop counts, or total travel time within traffic signal constraints.
    \item Problem formulation details:
    \begin{itemize}
        \item State variables: vehicle position and speed.
        \item Decision variables (actions): discrete acceleration set.
        \item Cost definition: incremental delay per timestep.
        \item Terminal constraint: vehicle must cross intersection during green signal window.
    \end{itemize}
\end{itemize}

\subsection{Algorithmic Approaches and Limitations}
\label{subsec:flow_dp_algorithms_limitations}
\begin{itemize}
    \item Common solution methods:
    \begin{itemize}
        \item Backward dynamic programming on spatial grid.
        \item A* search with heuristic pruning.
        \item Real-time optimization via receding horizon approaches.
        \item Co-optimization of signal timing (e.g., hierarchical DP).
    \end{itemize}
    \item Typical performance and limitations:
    \begin{itemize}
        \item Benchmark results: 20–50\,\% fewer stops, approx.\ 10\,\% travel-time savings.
        \item Sensitivity to prediction errors of arrival times.
        \item Ignorance of fuel efficiency, possible high-jerk trajectories.
        \item Computational complexity grows rapidly with finer discretizations and longer horizons.
    \end{itemize}
\end{itemize}