\section{Dynamic Programming for Energy-Aware GLOSA}
\label{sec:eco_glosa}

\subsection{Fuel Efficiency and Multi-Objective Optimization}
\label{subsec:eco_dp_objectives}
\begin{itemize}
    \item Extended optimization goals:
    \begin{itemize}
        \item Explicit minimization of fuel consumption and CO$_2$ emissions.
        \item Trade-offs between delay, fuel, and driver comfort (multi-objective Pareto optimization).
    \end{itemize}
    \item Common fuel and emission modeling approaches:
    \begin{itemize}
        \item Integration with detailed fuel/emission models (PHEMlight5, HBEFA4).
        \item Physics-based tractive-energy modeling.
    \end{itemize}
    \item Typical constraints in formulation:
    \begin{itemize}
        \item Vehicle dynamics constraints (speed, acceleration limits, legal speed boundaries).
        \item Powertrain limitations (maximum engine power, battery state-of-charge for hybrids).
    \end{itemize}
\end{itemize}

\subsection{Literature Results and Open Issues}
\label{subsec:eco_dp_results_issues}
\begin{itemize}
    \item Key studies and typical outcomes:
    \begin{itemize}
        \item Notable research: Khodayari et al.\ (2018, DP+EMIT), Ma et al.\ (2020, DP applied to Audi A3 e-tron).
        \item Additional fuel savings (3–7\,\%) over throughput-oriented DP at minor travel-time costs.
    \end{itemize}
    \item Remaining open issues and research gaps:
    \begin{itemize}
        \item Sensitivity to parameter uncertainty (engine characteristics, topography).
        \item Lack of studies on platoon-level interactions (egoistic vs.\ cooperative behaviors).
        \item Feasibility and computational demand for real-time applications in low-power embedded systems.
    \end{itemize}
\end{itemize}