\DeclareAcronym{acc}{
  short = ACC,
  long  = Adaptive Cruise Control,
  long-plural-form = Adaptive Cruise Control systems,
  extra = {A driver assistance system that automatically adjusts vehicle speed to maintain a safe distance from vehicles ahead.}
}

\DeclareAcronym{adas}{
  short = ADAS,
  long  = Advanced Driver-Assistance System,
  long-plural-form = Advanced Driver-Assistance Systems,
  extra = {Systems that assist drivers with driving and parking functions, designed to increase car safety and driver comfort.}
}

\DeclareAcronym{alks}{
  short = ALKS,
  long  = Automated Lane Keeping System,
  long-plural-form = Automated Lane Keeping Systems,
  extra = {A system that controls the lateral and longitudinal movement of a vehicle to keep it within its lane on highways.}
}

\DeclareAcronym{anova}{
  short = ANOVA,
  long  = Analysis of Variance,
  long-plural-form = Analyses of Variance,
  extra = {A statistical method used to test for significant differences between the means of two or more groups.}
}

\DeclareAcronym{asct}{
  short = ASCT,
  long  = Adaptive Signal Control Technology,
  long-plural-form = Adaptive Signal Control Technologies,
  extra = {Technology that optimizes traffic signal timing in real-time based on current traffic demand.}
}

\DeclareAcronym{asl}{
  short = ASL,
  long  = Advisory Speed Limit,
  long-plural-form = Advisory Speed Limits,
  extra = {Dynamically suggested speed recommendations to promote smoother traffic flow and enhance safety.}
}

\DeclareAcronym{bev}{
  short = BEV,
  long  = Battery Electric Vehicle,
  long-plural-form = Battery Electric Vehicles,
  extra = {A type of electric vehicle that exclusively uses chemical energy stored in rechargeable battery packs.}
}

\DeclareAcronym{c2x}{
  short = C2X,
  long  = Car-to-Everything,
  extra = {A communication framework enabling cars to interact with various external entities; often used interchangeably with V2X.}
}

\DeclareAcronym{cacc}{
  short = CACC,
  long  = Cooperative Adaptive Cruise Control,
  long-plural-form = Cooperative Adaptive Cruise Control systems,
  extra = {An extension of ACC that uses V2V communication to enable vehicles to travel in platoons with smaller headways.}
}

\DeclareAcronym{cav}{
  short = CAV,
  long  = Connected and Automated Vehicle,
  long-plural-form = Connected and Automated Vehicles,
  extra = {Vehicles equipped with both connectivity (V2X) and automation (ADAS) technologies.}
}

\DeclareAcronym{c-its}{
  short = C-ITS,
  long  = Cooperative Intelligent Transport System,
  long-plural-form = Cooperative Intelligent Transport Systems,
  extra = {Systems where vehicles and infrastructure share information to improve safety, traffic efficiency, and comfort.}
}

\DeclareAcronym{co2}{
  short = CO$_2$,
  long  = Carbon Dioxide,
  extra = {A primary greenhouse gas emitted through the combustion of fossil fuels.}
}

\DeclareAcronym{cop}{
  short = COP,
  long  = Control Optimization Problem,
  long-plural-form = Control Optimization Problems,
  extra = {A mathematical problem that involves finding a control policy for a given system such that a certain optimality criterion is achieved.}
}

\DeclareAcronym{cv}{
  short = CV,
  long  = Connected Vehicle,
  long-plural-form = Connected Vehicles,
  extra = {A vehicle capable of communicating with other systems outside of the vehicle, such as other vehicles or infrastructure.}
}

\DeclareAcronym{c-v2x}{
  short = C-V2X,
  long  = Cellular Vehicle-to-Everything,
  extra = {A communication technology using cellular networks to connect vehicles to each other and to roadside infrastructure.}
}

\DeclareAcronym{dddp}{
  short = DDDP,
  long  = Discrete Differential Dynamic Programming,
  long-plural-form = Discrete Differential Dynamic Programming methods,
  extra = {An optimization method for trajectory planning in discrete-time optimal control problems.}
}

\DeclareAcronym{dlr}{
  short = DLR,
  long  = German Aerospace Center,
  extra = {The national aeronautics and space research center of the Federal Republic of Germany.}
}

\DeclareAcronym{dp}{
  short = DP,
  long  = Dynamic Programming,
  extra = {A method for solving complex problems by breaking them down into simpler subproblems.}
}

\DeclareAcronym{ead}{
  short = EAD,
  long  = Eco-Approach and Departure,
  long-plural-form = Eco-Approach and Departure strategies,
  extra = {A driving strategy that uses traffic signal information to advise drivers on the most fuel-efficient speed profile when approaching and departing intersections.}
}

\DeclareAcronym{eco-glosa}{
  short = Eco-GLOSA,
  long  = Ecological Green Light Optimal Speed Advisory,
  long-plural-form = Ecological Green Light Optimal Speed Advisories,
  extra = {An extension of GLOSA that provides speed advice to minimize fuel consumption and emissions.}
}

\DeclareAcronym{eidm}{
  short = EIDM,
  long  = Extended Intelligent Driver Model,
  extra = {An enhanced car-following model that incorporates more realistic driver behaviors, such as reaction times and estimation errors.}
}

\DeclareAcronym{ev}{
  short = EV,
  long  = Electric Vehicle,
  long-plural-form = Electric Vehicles,
  extra = {A vehicle that is propelled by one or more electric motors, using energy stored in rechargeable batteries.}
}

\DeclareAcronym{fc}{
  short = FC,
  long  = Fuel Consumption,
  extra = {The amount of fuel a vehicle consumes to travel a certain distance.}
}

\DeclareAcronym{fcd}{
  short = FCD,
  long  = Floating Car Data,
  extra = {Data collected from vehicles that provides information on traffic flow, speed, and congestion.}
}

\DeclareAcronym{flow-glosa}{
  short = Flow-GLOSA,
  long  = Flow-Optimized Green Light Optimal Speed Advisory,
  long-plural-form = Flow-Optimized Green Light Optimal Speed Advisories,
  extra = {A GLOSA system that optimizes speed advice to improve overall traffic flow and reduce congestion.}
}

\DeclareAcronym{fot}{
  short = FOT,
  long  = Field Operational Test,
  long-plural-form = Field Operational Tests,
  extra = {Real-world studies of transport technologies to assess their impacts and user acceptance.}
}

\DeclareAcronym{ga}{
  short = GA,
  long  = Genetic Algorithm,
  long-plural-form = Genetic Algorithms,
  extra = {A search heuristic inspired by the process of natural selection, used for optimization and search problems.}
}

\DeclareAcronym{gan}{
  short = GAN,
  long  = Generative Adversarial Network,
  long-plural-form = Generative Adversarial Networks,
  extra = {A class of machine learning frameworks where two neural networks contest with each other in a zero-sum game.}
}

\DeclareAcronym{gdpr}{
  short = GDPR,
  long  = General Data Protection Regulation,
  extra = {A regulation in EU law on data protection and privacy for all individuals within the European Union and the European Economic Area.}
}

\DeclareAcronym{glosa}{
  short = GLOSA,
  long  = Green Light Optimal Speed Advisory,
  long-plural-form = Green Light Optimal Speed Advisories,
  extra = {A system that provides drivers with the optimal speed to approach a signalized intersection to pass on a green light.}
}

\DeclareAcronym{gui}{
  short = GUI,
  long  = Graphical User Interface,
  long-plural-form = Graphical User Interfaces,
  extra = {A visual interface that allows users to interact with electronic devices through graphical icons and visual indicators.}
}

\DeclareAcronym{hbefa}{
  short = HBEFA,
  long  = Handbook of Emission Factors for Road Transport,
  extra = {A database providing emission factors for all common road vehicle categories.}
}

\DeclareAcronym{hil}{
  short = HIL,
  long  = Hardware-in-the-Loop,
  long-plural-form = Hardware-in-the-Loop setups,
  extra = {A technique used for the development and testing of complex real-time embedded systems.}
}

\DeclareAcronym{hmi}{
  short = HMI,
  long  = Human-Machine Interface,
  long-plural-form = Human-Machine Interfaces,
  extra = {The user interface that connects an operator to a machine, system, or device.}
}

\DeclareAcronym{i2v}{
  short = I2V,
  long  = Infrastructure-to-Vehicle,
  extra = {Communication from roadside infrastructure to vehicles.}
}

\DeclareAcronym{icw}{
  short = ICW,
  long  = Intersection Collision Warning,
  long-plural-form = Intersection Collision Warning systems,
  extra = {A system designed to warn a driver of a potential collision at an intersection.}
}

\DeclareAcronym{idm}{
  short = IDM,
  long  = Intelligent Driver Model,
  long-plural-form = Intelligent Driver Models,
  extra = {A car-following model that describes the longitudinal motion of vehicles.}
}

\DeclareAcronym{its}{
  short = ITS,
  long  = Intelligent Transportation System,
  long-plural-form = Intelligent Transportation Systems,
  extra = {The application of sensing, analysis, control, and communications technologies to ground transportation to improve safety, mobility, and efficiency.}
}

\DeclareAcronym{mae}{
  short = MAE,
  long  = Mean Absolute Error,
  extra = {A measure of errors between paired observations expressing the same phenomenon.}
}

\DeclareAcronym{map}{
  short = MAP,
  long  = Map Data Message,
  long-plural-form = Map Data Messages,
  extra = {A V2X message that provides detailed information about the geometry and topology of road intersections.}
}

\DeclareAcronym{marl}{
  short = MARL,
  long  = Multi-Agent Reinforcement Learning,
  extra = {A subfield of reinforcement learning where multiple agents learn concurrently in a shared environment.}
}

\DeclareAcronym{mdp}{
  short = MDP,
  long  = Markov Decision Process,
  long-plural-form = Markov Decision Processes,
  extra = {A mathematical framework for modeling decision-making in situations where outcomes are partly random and partly under the control of a decision-maker.}
}

\DeclareAcronym{milp}{
  short = MILP,
  long  = Mixed-Integer Linear Programming,
  extra = {An optimization problem in which some of the variables are constrained to be integers, while others are continuous.}
}

\DeclareAcronym{mpc}{
  short = MPC,
  long  = Model Predictive Control,
  long-plural-form = Model Predictive Control schemes,
  extra = {An advanced method of process control that is used to control a process while satisfying a set of constraints.}
}

\DeclareAcronym{mpg}{
  short = MPG,
  long  = Miles Per Gallon,
  extra = {A measure of fuel economy, representing the number of miles a vehicle can travel per gallon of fuel.}
}

\DeclareAcronym{mpr}{
  short = MPR,
  long  = Market Penetration Rate,
  long-plural-form = Market Penetration Rates,
  extra = {The percentage of vehicles in a traffic stream that are equipped with a specific technology.}
}

\DeclareAcronym{nox}{
  short = NO$_x$,
  long  = Nitrogen Oxides,
  extra = {A generic term for the various nitrogen oxides produced during combustion.}
}

\DeclareAcronym{np}{
  short = NP,
  long  = Non-deterministic Polynomial time,
  extra = {A complexity class used to classify decision problems.}
}

\DeclareAcronym{obu}{
  short = OBU,
  long  = On-Board Unit,
  long-plural-form = On-Board Units,
  extra = {A device installed in a vehicle that enables V2X communication.}
}

\DeclareAcronym{ode}{
  short = ODE,
  long  = Ordinary Differential Equation,
  long-plural-form = Ordinary Differential Equations,
  extra = {A differential equation containing one or more functions of one independent variable and its derivatives.}
}

\DeclareAcronym{oem}{
  short = OEM,
  long  = Original Equipment Manufacturer,
  long-plural-form = Original Equipment Manufacturers,
  extra = {A company that produces parts and equipment that may be marketed by another manufacturer.}
}

\DeclareAcronym{ovm}{
  short = OVM,
  long  = Optimal Velocity Model,
  long-plural-form = Optimal Velocity Models,
  extra = {A car-following model that determines the optimal velocity of a vehicle based on the distance to the preceding vehicle.}
}

\DeclareAcronym{phem}{
  short = PHEM,
  long  = Passenger Car and Heavy-Duty Emission Model,
  extra = {A simulation tool for calculating fuel consumption and emissions of road vehicles.}
}

\DeclareAcronym{pm}{
  short = PM,
  long  = Particulate Matter,
  extra = {The sum of all solid and liquid particles suspended in air, many of which are hazardous.}
}

\DeclareAcronym{rl}{
  short = RL,
  long  = Reinforcement Learning,
  extra = {An area of machine learning concerned with how intelligent agents ought to take actions in an environment to maximize the notion of cumulative reward.}
}

\DeclareAcronym{rmse}{
  short = RMSE,
  long  = Root Mean Square Error,
  extra = {A frequently used measure of the differences between values predicted by a model and the values observed.}
}

\DeclareAcronym{rq}{
  short = RQ,
  long  = Research Question,
  long-plural-form = Research Questions,
  extra = {A question that a research project sets out to answer.}
}

\DeclareAcronym{rsu}{
  short = RSU,
  long  = Roadside Unit,
  long-plural-form = Roadside Units,
  extra = {A piece of infrastructure that communicates with On-Board Units in vehicles to provide V2I services.}
}

\DeclareAcronym{sdp}{
  short = SDP,
  long  = Stochastic Dynamic Programming,
  extra = {A framework for modeling and solving sequential decision-making problems under uncertainty.}
}

\DeclareAcronym{spat}{
  short = SPaT,
  long  = Signal Phase and Timing,
  extra = {Data that describes the current status and future timing of traffic signal phases.}
}

\DeclareAcronym{sumo}{
  short = SUMO,
  long  = Simulation of Urban MObility,
  extra = {An open-source, microscopic, multi-modal traffic simulation suite.}
}

\DeclareAcronym{tmc}{
  short = TMC,
  long  = Traffic Management Center,
  long-plural-form = Traffic Management Centers,
  extra = {A facility that monitors and controls traffic on a road network.}
}

\DeclareAcronym{traci}{
  short = TraCI,
  long  = Traffic Control Interface,
  extra = {An API for coupling traffic simulations with external applications.}
}

\DeclareAcronym{ttg}{
  short = TTG,
  long  = Time-to-Green,
  extra = {The time remaining until a traffic light turns green.}
}

\DeclareAcronym{ttr}{
  short = TTR,
  long  = Time-to-Red,
  extra = {The time remaining until a traffic light turns red.}
}

\DeclareAcronym{v2i}{
  short = V2I,
  long  = Vehicle-to-Infrastructure,
  extra = {Communication between vehicles and roadside infrastructure.}
}

\DeclareAcronym{v2v}{
  short = V2V,
  long  = Vehicle-to-Vehicle,
  extra = {Communication between vehicles.}
}

\DeclareAcronym{v2x}{
  short = V2X,
  long  = Vehicle-to-Everything,
  extra = {The umbrella term for a vehicle's communication system, where information is passed from a vehicle to any entity that may affect the vehicle, and vice versa.}
}

\DeclareAcronym{ecoacc}{
  short             = Eco-ACC,
  long              = Eco-Cooperative Adaptive Cruise Control,
  short-plural-form = Eco-ACC systems,
  long-plural-form  = Eco-Cooperative Adaptive Cruise Control systems,
  extra             = {An enhancement of adaptive cruise control that optimises fuel efficiency and emissions by cooperating with traffic signal information and surrounding vehicle data.}
}

\DeclareAcronym{obd}{
  short             = OBD,
  long              = On-Board Diagnostics,
  short-plural-form = OBD systems,
  long-plural-form  = On-Board Diagnostics systems,
  extra             = {An automotive term referring to a vehicle’s self-diagnostic and reporting capability, used to monitor and analyze engine and component performance.}
}

\DeclareAcronym{tli}{
  short             = TLI,
  long              = Traffic Light Information,
  short-plural-form = TLI messages,
  long-plural-form  = Traffic Light Information messages,
  extra             = {A data exchange service that provides real-time signal phase and timing (SPaT) and map data to connected vehicles for safety and eco-driving applications.}
}

\DeclareAcronym{td}{
  short               = TD,
  long                = Temporal-Difference Reinforcement Learning,
  short-plural-form   = TD methods,
  long-plural-form    = Temporal-Difference Reinforcement Learning methods,
  extra               = {A class of reinforcement learning algorithms that update estimates based on the difference between predicted and actual rewards over time.}
}

\DeclareAcronym{adp}{
  short               = ADP,
  long                = Approximate Dynamic Programming,
  short-plural-form   = ADP methods,
  long-plural-form    = Approximate Dynamic Programming methods,
  extra               = {A set of techniques for solving dynamic programming problems approximately when exact solutions are computationally intractable.}
}

\DeclareAcronym{pl}{
  short               = PL,
  long                = Numerical Perturbation Learning,
  short-plural-form   = PL methods,
  long-plural-form    = Numerical Perturbation Learning methods,
  extra               = {A learning approach that estimates gradients or policy updates via small numerical perturbations to the system.}
}

\DeclareAcronym{sh}{
  short               = SH,
  long                = Shooting Heuristic,
  short-plural-form   = SHs,
  long-plural-form    = Shooting Heuristics,
  extra               = {A simplified optimization approach that analytically estimates vehicle trajectories by \enquote{shooting} from current states toward desired conditions.}
}

\DeclareAcronym{ttc}{
  short               = TTC,
  long                = Time-to-Cross,
  short-plural-form   = TTCs,
  long-plural-form    = Time-to-Cross values,
  extra               = {Estimated time required for a vehicle to reach and traverse an intersection or control point.}
}

\DeclareAcronym{vgdp}{
  short               = VGDP,
  long                = Variable-Grid Dynamic Programming,
  short-plural-form   = VGDPs,
  long-plural-form    = Variable-Grid Dynamic Programming methods,
  extra               = {A dynamic programming variant that uses non-uniform discretisation to reduce computational cost while preserving solution accuracy.}
}

\DeclareAcronym{msdp}{
  short               = MS-DP,
  long                = Multi-Stage Dynamic Programming,
  short-plural-form   = MS-DPs,
  long-plural-form    = Multi-Stage Dynamic Programming methods,
  extra               = {A dynamic programming approach that decomposes complex control problems into sequential sub-problems for improved tractability.}
}

\DeclareAcronym{qost}{
  short               = QOST,
  long                = Queue-Aware Optimal Speed Trajectory,
  short-plural-form  = QOSTs,
  long-plural-form   = Queue-Aware Optimal Speed Trajectories,
  extra              = {A trajectory planning strategy that accounts for upstream queues when determining fuel- or delay-optimal speed profiles.}
}

\DeclareAcronym{ecomsq}{
  short               = Eco-MS-Q,
  long                = Modular Queue-Aware Multi-Signal Energy-Optimal Controller,
  short-plural-form  = Eco-MS-Q controllers,
  long-plural-form   = Modular Queue-Aware Multi-Signal Energy-Optimal Controllers,
  extra               = {A control architecture integrating queue detection and multi-phase signal coordination for energy-optimal trajectory planning.}
}

\DeclareAcronym{soc}{
  short               = SoC,
  long                = State of Charge,
  short-plural-form   = SoC levels,
  long-plural-form    = States of Charge,
  extra               = {The available battery energy expressed as a percentage of total capacity.}
}
